---
title: "Day 21 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


## Part 1
Because there are a limited number of combinations - we'll just expand them all out.
```{r}
input<-read_lines("../../AocData/AOC2015/Day21.txt")
blanklines<-which(input=="")
```
Start by creating 3 data frames - one for each thing:
```{r}
### weapons
weapons<-as.data.frame(matrix(ncol=4,nrow=0))
for(i in 2:(blanklines[1]-1)){
  a<-input[i]
  a<-unlist(str_split(str_squish(a)," "))
  weapons<-rbind(weapons,a)}
colnames(weapons)<-c("weapon","wcost","wdamage","wdefense")
weapons$wcost<-as.numeric(weapons$wcost)
weapons$wdamage<-as.numeric(weapons$wdamage)
weapons$wdefense<-as.numeric(weapons$wdefense)

### armor
armor<-as.data.frame(list("no armor",0,0,0))
for(i in (blanklines[1]+2):(blanklines[2]-1)){
  a<-input[i]
  a<-unlist(str_split(str_squish(a)," "))
  armor<-rbind(armor,a)}
colnames(armor)<-c("armor","acost","adamage","adefense")
armor$acost<-as.numeric(armor$acost)
armor$adamage<-as.numeric(armor$adamage)
armor$adefense<-as.numeric(armor$adefense)


ring<-as.data.frame(list("no ring",0,0,0))
for(i in (blanklines[2]+2):(length(input))){
  a<-input[i]
  a<-str_replace(a," ","")
  a<-unlist(str_split(str_squish(a)," "))
  ring<-rbind(ring,a)}

rings<-as.data.frame(list("no ring",0,0,0,"no ring",0,0,0))
for(i in 1:(nrow(ring)-1)){
  for(j in (i+1):nrow(ring)){
    x<-unname(append(ring[i,],ring[j,]))
    rings<-rbind(rings,x)}}
colnames(rings)<-c("ring1","r1cost","r1damage","r1defense","ring2","r2cost","r2damage","r2defense")
rings$r1cost<-as.numeric(rings$r1cost)
rings$r2cost<-as.numeric(rings$r2cost)
rings$r1damage<-as.numeric(rings$r1damage)
rings$r2damage<-as.numeric(rings$r2damage)
rings$r1defense<-as.numeric(rings$r1defense)
rings$r2defense<-as.numeric(rings$r2defense)


```

Then combine them into one dataframe
```{r}
outfit<-as.data.frame(matrix(nrow=0,ncol=16))
for(w in 1:nrow(weapons)){
  for(a in 1:nrow(armor)){
    for(r in 1:nrow(rings)){
      war<-cbind(weapons[w,],armor[a,],rings[r,])
      outfit<-rbind(outfit,war)}}}

```

So - now calculate all the stuff that matters:  total cost, damage points, armor points
```{r}
outfit <- outfit %>% rowwise %>% mutate(totalcost=r1cost+r2cost+wcost+acost,
                                        damagepoints=wdamage+adamage+r1damage+r2damage,
                                        defensepoints=wdefense+adefense+r1defense+r2defense)
```

Then, find the points per turn & the number of turns to lose & whether or not you win
```{r}
bossdamage<-7
bossarmor<-2
bosshitpoints<-12
playerhitpoints<-10
outfit <- outfit %>% rowwise %>%
  mutate(playerdamage=max(bossdamage-defensepoints,1),
         bossdamage=(max(damagepoints-bossarmor,1))) %>%
  ### death is the number of whole turns before hitting 0, 100 is player hit points
  mutate(playerdeathat=ceiling(playerhitpoints/playerdamage),
         bossdeathat=ceiling(bosshitpoints/bossdamage)) %>%
  ### because the player goes first, if it takes more (or the same) turns, the player wins
  mutate(win=ifelse(playerdeathat>=bossdeathat,"Win","Lose"))
### sample outfits
select(outfit,c(weapon,armor,ring1,ring2,damagepoints,defensepoints,totalcost,win))[sample(1:nrow(outfit),8),]
```


```{r}
cheapest<- outfit %>% group_by(win) %>% summarise(tc=min(totalcost))
part1<-cheapest%>%filter(win=="Win")%>%select(tc)
part1
```

## Part 2
Everything is already set up, just find out which is the most expensive loss
```{r}
costliest <- outfit %>% group_by(win) %>% summarize(tc=max(totalcost))
part2 <- costliest %>% filter(win=="Lose") %>% select(tc)
part2
```
