---
title: "Day 24 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(numbers)
options(scipen = 999)
```

```{r}
input<-c(1:5,7:11)
pcks<-as.numeric(input)
```


## Part 1
Taking a gamble - hopefully, all of the small groups will be "valid" (i.e., there is a way to sort the packages into groups.)  This isn't certain, but likely.  If so - then, the smallest product of the smallest groups will be the answer
```{r}
### sort from largest to smallest, to keep the space narrow.
pcks<-sort(pcks,decreasing=TRUE)
### target is 1/3rd of the packs
target<-sum(pcks/3)
### create a working data frame
df<-data.frame(c(list(0,sum(pcks),0,1)))
colnames(df)<-c("total","possible","packsize","QE")
### tiny pack is the size of the smallest pack found so far
tinypack<-length(pcks)
### hits is the group of valid packs (adds to the target and all smallest found so far)
hits<-as.data.frame(matrix(ncol=4,nrow=0))
for(i in 1:length(pcks)){
  ### total of how many have *not* been sorted yet.
  rmn<-sum(pcks[i:length(pcks)])-pcks[i]
  ncname<-c(colnames(df),paste0("P",pcks[i]))
  ### copy the dataframe - add the first element of the list to one copy, don't to the other.
  ### keep track of the total for each pack, the possible total for each pack, the QE and the number of packages
  with<-df %>% rowwise %>%
    mutate(total=total+pcks[i]) %>%
    mutate(possible=rmn+total) %>%
    mutate(packsize=packsize+1)%>%
    mutate(QE=QE*pcks[i])%>%
    mutate(x=pcks[i])
  without<-df %>% rowwise %>%
    mutate(possible=rmn+total,x=0)
  df<-rbind(with,without)
  ### make sure that recombining works
  colnames(df)<-ncname
  hits<-hits%>%mutate(x=0)
  colnames(hits)<-ncname
  ### package groups that meet the total to the list of valid hits
  newhits<-df %>% filter(total==target)
  hits<-rbind(hits,newhits)%>%ungroup%>%filter(packsize<=tinypack)
  ### filter out anything that is too big, will never get big enough, or already has too many packages
  df<-df %>% filter(total<target,possible>=target,packsize<=tinypack)
  ### reset the max number of packages
  tinypack<-min(hits$packsize,tinypack)}
### again, filter out all but the smallest groups
hits<-hits%>%ungroup%>%filter(packsize<=tinypack)
hits
```
hits should be the all of the smallest possible size packages.  Because I BELIEVE that there must be some way of forming 3 packs with the smallest - the smallest QE should be the answer
```{r}
part1<-min(hits$QE)
part1
```
## Part 2
Same thing, target is 4 groups - 

```{r}
target<-sum(pcks/4)
df<-data.frame(c(list(0,sum(pcks),0,1)))
colnames(df)<-c("total","possible","packsize","QE")
tinypack<-length(pcks)
hits<-as.data.frame(matrix(ncol=4,nrow=0))
for(i in 1:length(pcks)){
  rmn<-sum(pcks[i:length(pcks)])-pcks[i]
  ncname<-c(colnames(df),paste0("P",pcks[i]))
  with<-df %>% rowwise %>%
    mutate(total=total+pcks[i]) %>%
    mutate(possible=rmn+total) %>%
    mutate(packsize=packsize+1)%>%
    mutate(QE=QE*pcks[i])%>%
    mutate(x=pcks[i])
  without<-df %>% rowwise %>%
    mutate(possible=rmn+total,x=0)
  df<-rbind(with,without)
  colnames(df)<-ncname
  hits<-hits%>%mutate(x=0)
  colnames(hits)<-ncname
  newhits<-df %>% filter(total==target)
  hits<-rbind(hits,newhits)%>%ungroup%>%filter(packsize<=tinypack)
  df<-df %>% filter(total<target,possible>=target,packsize<=tinypack)
  tinypack<-min(hits$packsize,tinypack)}
hits<-hits%>%ungroup%>%filter(packsize<=tinypack)
hits
```


```{r}
part2<-min(hits$QE)
part2
```