---
title: "Day 18 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


## Part 1

Change the grid to a matrix - but to make things less complex, the matrix will be 0's and 1's (for off and on) and be slightly bigger with a row of 0's around the edge


```{r}
input<-read_lines("Day18Sample.txt")

lightgrid<-matrix(NA,nrow=2+length(input),ncol=2+nchar(input[1]))
lightgrid[1,]<-0
lightgrid[,1]<-0
lightgrid[nrow(lightgrid),]<-0
lightgrid[,ncol(lightgrid)]<-0
for(i in 1:length(input)){
  x<-input[i]
  x<-str_replace_all(x,"\\.","0")
  x<-str_replace_all(x,"#","1")
  x<-as.numeric(unlist(str_split(x,"")))
  lightgrid[i+1,2:(ncol(lightgrid)-1)]<-x}
```

Quick function that takes in a matrix and returns the animated matrix:

```{r}
animate<-function(x=lightgrid){
  newgrid<-matrix(ncol=ncol(x),nrow=nrow(x))
  newgrid[1,]<-0
  newgrid[,1]<-0
  newgrid[nrow(newgrid),]<-0
  newgrid[,ncol(newgrid)]<-0
  for(i in 2:(ncol(newgrid)-1)){
    for(j in 2:(nrow(newgrid)-1)){
      ctr<-x[i,j]
      rng<-sum(x[(i-1):(i+1),(j-1):(j+1)])-ctr
      if(ctr==0){
        if(rng==3){newgrid[i,j]<-1}else{newgrid[i,j]<-0}
      }else{if(rng==2||rng==3){newgrid[i,j]<-1}else{newgrid[i,j]<-0}}
      }}
  newgrid}
```
then another to repeat the animations:

```{r}
repeatanimate<-function(x=lightgrid,reps=100){
  a<-x
  for(i in 1:reps){
    a<-animate(a)}
  a}
```
Part1 is the sum of the repeated animation
```{r}
part1<-repeatanimate()
part1<-sum(part1)
part1
```
## Part 2
With the corners stuck, what happens?
```{r}
stuck<-function(x=lightgrid){
  newgrid<-matrix(ncol=ncol(x),nrow=nrow(x))
  newgrid[1,]<-0
  newgrid[,1]<-0
  newgrid[nrow(newgrid),]<-0
  newgrid[,ncol(newgrid)]<-0
  for(i in 2:(ncol(newgrid)-1)){
    for(j in 2:(nrow(newgrid)-1)){
      ctr<-x[i,j]
      rng<-sum(x[(i-1):(i+1),(j-1):(j+1)])-ctr
      if(ctr==0){
        if(rng==3){newgrid[i,j]<-1}else{newgrid[i,j]<-0}
      }else{if(rng==2||rng==3){newgrid[i,j]<-1}else{newgrid[i,j]<-0}}
    }}
  ###change the corners to 1, no matter what
  newgrid[2,2]<-1
  newgrid[2,(nrow(newgrid)-1)]<-1
  newgrid[(ncol(newgrid)-1),2]<-1
  newgrid[(ncol(newgrid)-1),(nrow(newgrid)-1)]<-1
  newgrid}
```

```{r}
stuckgrid<-lightgrid
  stuckgrid[2,2]<-1
  stuckgrid[2,(nrow(stuckgrid)-1)]<-1
  stuckgrid[(ncol(stuckgrid)-1),2]<-1
  stuckgrid[(ncol(stuckgrid)-1),(nrow(stuckgrid)-1)]<-1

```
and a new repeating function
```{r}
stuckrepeat<-function(x=stuckgrid,reps=100){
  a<-x
  for(i in 1:reps){
    a<-stuck(a)}
  a}
```
Part 2 is the sum of the stuck animation
```{r}
part2<-sum(stuckrepeat())
part2
```

