---
title: "Day 12 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


## Part 1
For part 1, find all the numbers in the document and add them together:

```{r}
input<-read_lines("Day12Sample.txt")
nums<-c("-","0","1","2","3","4","5","6","7","8","9") ## include all digits and the negative sign
### replace everything that isn't a number or a negative sign with a space
part1<-str_replace_all(input,"[^0123456789-]"," ")
### remove the extra spaces, turn it into a numeric vector and add it up
part1<-sum(as.numeric(unlist(str_split(str_squish(part1)," "))))
part1
```

## Part 2
I should know more about JSON than I do.  But I don't, so not going to use it.  Instead, identify things with the pattern *:"red"* inside curly brackets and then delete everything within those curly brackets.

```{r}
### Create a function "pullobject"
pullobject<-function(st){
  #  if there are any left curly brackets,
  while(str_count(st,"\\{")>0){
    ### find the position of the first right bracket
    firstrightbracket<-str_locate(st,"\\}")[1,2]
    ### find all the left brackets up to that point and pick the furthest right of those
    closestleftbracket<-tail(str_locate_all(str_sub(st,1,firstrightbracket),"\\{")[[1]],1)[,2]
    ### if it has the a :"red" inside (property "red") delete from bracket to bracket
    if(str_detect(str_sub(st,closestleftbracket,firstrightbracket),"\":\"red\"")){
      st<-str_c(str_sub(st,1,closestleftbracket-1),str_sub(st,firstrightbracket+1,nchar(st)))
    }else{
      ### if not, change the brackets to greater than and less than signs
      str_sub(st,firstrightbracket,firstrightbracket)<-">"
      str_sub(st,closestleftbracket,closestleftbracket)<-"<"}}
  ### return the altered string
  st}
```


Run input on through pullobject to get a new string and add the numbers.   This does not work on the sample data.
```{r,eval=FALSE}
part2<-pullobject(input)
part2<-str_replace_all(part2,"[^0123456789-]"," ")
part2<-sum(as.numeric(unlist(str_split(str_squish(part2)," "))))
part2
```


