---
title: "Day 7 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
library(collections)
options(scipen = 999)
```


## Part 1
```{r}
input <- read_lines("Day7Sample.txt")
```

I think that this might make sense as a dictionary - store the values and then add, subtract, etc. as needed.
Create the dictionary, then create the function that handles each of the types of instructions with the dictionary
```{r}
circuitkit<-collections::dict()

runinstruction<-function(inst,d=circuitkit){
  ###split the instructions into pieces.
  ###need to handle AND, OR, LSHIFT, RSHIFT, NOT
  inst<-unlist(str_split(inst," "))
  if("AND" %in% inst){
    ### check to see if numbers
    if(str_detect(inst[1],"[:alpha:]",negate=TRUE)&&d$has(inst[3])){
      d$set(inst[5],bitwAnd(as.numeric(inst[1]),d$get(inst[3])))}else{
    if(d$has(inst[1])&&d$has(inst[3])){
      d$set(inst[5],bitwAnd(d$get(inst[1]),d$get(inst[3])))}}
  }else if("OR" %in% inst){
    if(d$has(inst[1])&&d$has(inst[3])){
      d$set(inst[5],bitwOr(d$get(inst[1]),d$get(inst[3])))}
  }else if("NOT" %in% inst){
    if(d$has(inst[2])){
      d$set(inst[4],bitwNot(d$get(inst[2])))}
  }else if("LSHIFT" %in% inst){
    if(d$has(inst[1])){
      d$set(inst[5],bitwShiftL(d$get(inst[1]),inst[3]))}
  }else if("RSHIFT" %in% inst){
    if(d$has(inst[1])){
      d$set(inst[5],bitwShiftR(d$get(inst[1]),inst[3]))}
  }else{
    if(str_detect(inst[1],"[:alpha:]",negate=TRUE)){
    d$set(inst[3],as.numeric(inst[1]))}else{
      if(d$has(inst[1])){
      d$set(inst[3],d$get(inst[1]))}}}
  d}
```


This is actually very static, the values (once set) do not change - but it does take some time to set & reset them.  So the instructions run over and over and over as it is slowly populated.

As soon as "a" exists, exit and that's part 1.
```{r}
i<-0
while(!circuitkit$has("a")&&i<1000){
for(i in 1:length(input)){
  circuitkit<-runinstruction(input[i],circuitkit)}
  i<-i+1}

part1<-circuitkit$get("a")
part1
```
## Part 2

For part 2, set b to the value from a & start over with re-running the instructions (skipping the instruction that would give a value to b.)

```{r}
### reset the kit
circuitkit<-collections::dict()

### set b to a
circuitkit$set("b",part1)

i<-0
while(!circuitkit$has("a")&&i<1000){
  for(i in 1:length(input)){
    ### don't do the instruction that resets b
    if(str_sub(input[i],-2)!=" b"){
      circuitkit<-runinstruction(input[i],circuitkit)}}
  i<-i+1}

part2<-circuitkit$get("a")
part2
```
```{r}
day7answers <-function(input){
  circuitkit<-collections::dict()
while(!circuitkit$has("a")){
for(i in 1:length(input)){
  circuitkit<-runinstruction(input[i],circuitkit)}}

part1<-circuitkit$get("a")
### reset the kit
circuitkit<-collections::dict()

### set b to a
circuitkit$set("b",part1)

while(!circuitkit$has("a")){
  for(i in 1:length(input)){
    ### don't do the instruction that resets b
    if(str_sub(input[i],-2)!=" b"){
      circuitkit<-runinstruction(input[i],circuitkit)}}}

part2<-circuitkit$get("a")
c(part1,part2)}

```

```{r}
day7 <- day7answers(input)
day7
```